Supplement to Global Distribution of Unintended Pregnancies and Contraceptive Usage Project

### 2024-04-19 code/eda-code/eda.qmd
Removed distribution graphs from manuscript. Edited region graph

# explore percent unintended pregnancies using short acting reversible methods by region
plot_4 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard, color = region)) + 
  geom_point() + 
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
plot(plot_4)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png")
ggsave(filename = figure_file, plot=plot_4)

@@

# explore the current marriage predictor
plot_7 <- mydata %>% ggplot(aes(x=pct_currentlymarried)) + 
  geom_histogram() +
  ggtitle("Percent of Women Currently Married") + 
  xlab("Percentage") + ylab("Number of Countries")
plot(plot_7)
figure_file = here("results", "figures", "exploratory-figures", "currmarried_distribution.png")
ggsave(filename = figure_file, plot=p23)

>>>>

# explore percent unintended pregnancies using short acting reversible methods by region
plot_4 <- mydata %>% ggplot(aes(x=pct_upreg_sarc, y=sarc_standard)) + 
  geom_point() + 
  geom_jitter(aes(colour = region)) +
  geom_smooth(method='lm') + 
  ggtitle("Percent of Unintended Pregnancies vs Standardzied Count of Women Using SARCs") + xlab("Standarized Total Number of Women") + ylab("Percentage of Unintended Pregnancies")
plot(plot_4)
figure_file = here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png")
ggsave(filename = figure_file, plot=plot_4)

@@

# explore the current marriage predictor
plot_7 <- mydata %>% ggplot(aes(x=pct_currentlymarried)) + 
  geom_histogram() +
  ggtitle("Percent of Women Currently Married") + 
  xlab("Percentage") + ylab("Number of Countries")
plot(plot_7)
figure_file = here("results", "figures", "exploratory-figures", "currmarried_distribution.png")
ggsave(filename = figure_file, plot=plot_7)

### 2024-04-19 products/manuscript/Manuscript.qmd
Removed distribution graphs from manuscript. Edited region graph

Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @fig-5 - @fig-10. The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes about 35 countries. Latin America/the Caribbean accounts for about 25 countries. Europe and Oceania account for the least number of observations. The distribution of women currently married is approximately normal with an average around 57% and no major outliers. The distribution of women never married is slightly skewed to the right with an average around 30%, but there are a few outliers greater than 60%. The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers. We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality, which each belong to the African region. The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram. Chad and South Sudan also account for the highest rates of maternal mortality in this category. The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor, which both belong to the African region.
```{r}
#| message: false
#| label: fig-5 - fig 10
#| fig-cap: "insert explanation here"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "region_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "currmarried_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "nevmarried_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeaths_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png"))

>>>>

Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @fig-5 - @fig-10. The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes about 35 countries. Latin America/the Caribbean accounts for about 25 countries. Europe and Oceania account for the least number of observations. The distribution of women currently married is approximately normal with an average around 57% and no major outliers. The distribution of women never married is slightly skewed to the right with an average around 30%, but there are a few outliers greater than 60%. The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers. We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality, which each belong to the African region. The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram. Chad and South Sudan also account for the highest rates of maternal mortality in this category. The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor, which both belong to the African region.

### 2024-04-19 products/manuscript/supplement/Supplementary-Material.qmd
Added labels to figures in manuscript, switched axis labels.
Removed distribution graphs from manuscript. Edited region graph

# Overview
This file provides extra information that is not included in our manuscript or the main processing and exploratory data analysis files. This information provides guidance on our thought processes behind the scenes and explains why we chose a single outcome to study.

@@

>>>>

# Overview
This file provides extra information that is not included in our manuscript or the main processing and exploratory data analysis files. This information provides guidance on our thought processes behind the scenes and explains why we chose a single outcome to study. The plots below are numbered based off of order of creation, they are not numbered in order.

@@

We also needed to explore the remaining predictor variables: `region`, `pct_currentlymarried`, `pct_nevermarried`, `rate_matdeaths`, `rate_matdeaths_upreg`, and `rate_matdeaths_abs`. Simple histograms will allow us to see the distribution of each variable and ensure there are no outliers. 
```{r}
# explore the region predictor
plot_6 <- mydata %>% ggplot(aes(x=region)) + 
  geom_bar() +
  ggtitle("Countries in Each Region") + 
  xlab("Region") + ylab("Number of Countries")
plot(plot_6)
figure_file = here("results", "figures", "exploratory-figures", "region_distribution.png")
ggsave(filename = figure_file, plot=plot_6)

# explore the current marriage predictor
plot_7 <- mydata %>% ggplot(aes(x=pct_currentlymarried)) + 
  geom_histogram() +
  ggtitle("Percent of Women Currently Married") + 
  xlab("Percentage") + ylab("Number of Countries")
plot(plot_7)
figure_file = here("results", "figures", "exploratory-figures", "currmarried_distribution.png")
ggsave(filename = figure_file, plot=plot_7)

# explore the never married predictor
plot_8 <- mydata %>% ggplot(aes(x=pct_nevermarried)) + 
  geom_histogram() +
  ggtitle("Percent of Women Never Married") + 
  xlab("Percentage") + ylab("Number of Countries")
plot(plot_8)
figure_file = here("results", "figures", "exploratory-figures", "nevmarried_distribution.png")
ggsave(filename = figure_file, plot=plot_8)

# explore the total maternal mortality rate predictor
plot_9 <- mydata %>% ggplot(aes(x=rate_matdeaths)) + 
  geom_histogram() +
  ggtitle("Total Rate of Maternal Deaths") + 
  xlab("Maternal Deaths") + ylab("Number of Countries")
plot(plot_9)
figure_file = here("results", "figures", "exploratory-figures", "matdeaths_distribution.png")
ggsave(filename = figure_file, plot=plot_9)

# explore outliers in maternal mortality from unintended pregnancies
mydata %>% filter(rate_matdeaths > 1000)

# explore the maternal mortality rate from unintended pregnancies predictor
plot_10 <- mydata %>% ggplot(aes(x=rate_matdeaths_upreg)) + 
  geom_histogram() +
  ggtitle("Rate of Maternal Deaths from Unintended Pregnancies") + 
  xlab("Maternal Deaths") + ylab("Number of Countries")
plot(plot_10)
figure_file = here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png")
ggsave(filename = figure_file, plot=plot_10)

# explore outliers in maternal mortality from unintended pregnancies
mydata %>% filter(rate_matdeaths_upreg > 6000)

# explore the maternal mortality rate from abortions
plot_11 <- mydata %>% ggplot(aes(x=rate_matdeaths_abs)) + 
  geom_histogram() +
  ggtitle("Rate of Maternal Deaths from Abortions") + 
  xlab("Maternal Deaths") + ylab("Number of Countries")
plot(plot_11)
figure_file = here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png")
ggsave(filename = figure_file, plot=plot_11)

# explore outliers in maternal mortality from abortions
mydata %>% filter(rate_matdeaths_abs > 500)
